 Reproducable Research
======================

###Peer Assessment 1

####Loading and preprocessing the data


```{r, echo=TRUE}
activity<-read.csv("activity.csv", header=T,colClasses=c("numeric","Date","numeric"))
```

####Mean total number of steps taken per day  

Let's compute the total number of steps taken per day:

```{r, echo=TRUE}
totalstepsperday<-tapply(activity$steps,activity$date,sum)
```
and make histogram

```{r echo=TRUE}
hist(totalstepsperday,main="Histogram of the total number of steps taken each day",xlab="Number of steps",col="grey")

```

Calculate and report the mean and median total number of steps taken
per day 

```{r, echo=TRUE}
mean_steps<- mean(totalstepsperday, na.rm=T)
median_steps<-median(totalstepsperday, na.rm=T)
```

Mean steps taken per day is **`r mean_steps`**

Median steps taken per day is **`r median_steps`**


####Average daily activity pattern

Making a time series plot of the average number of steps taken within each 5-minute interval (averaged across all days)

```{r echo=TRUE}

int_average<-tapply(activity$steps,activity$interval,mean,na.rm=T)
plot(names(int_average),int_average,type="l",xlab="Interval",ylab="Average number of steps")
```

Let's find 5-minute interval that contains the maximum number of steps on average across all the days.

```{r echo=TRUE}
maxsteps<-max(int_average)
int_max<-names(int_average[int_average==maxsteps])
```

So, this is the interval **`r int_max`** and maximum number of steps is **`r maxsteps`**

####Imputing missing values

Calculate and report the total number of rows with missing values 

```{r echo=TRUE}
miss_val_rows<-nrow(activity)-sum(complete.cases(activity))
```

Total number of rows with missing values is **`r miss_val_rows`**  

Filling in all of the missing values in the dataset

The stategy is to substitute missing values using mean of corresponding 5-minute interval

Creating a new dataset (named activity1) with the missing data filled in

```{r echo=TRUE}
#int_average is the vector of means for each interval and was previously calculated
activity1<-activity
missed_names<-as.character(activity1[is.na(activity1[,"steps"]),"interval"])
missed_values<-int_average[missed_names]

activity1[is.na(activity1[,"steps"]),"steps"]<-missed_values
```

Histogram of the total number of steps taken each day with the missing data filled in

```{r, echo=TRUE}
totalstepsperday1<-tapply(activity1$steps,activity1$date,sum)
hist(totalstepsperday1,main="Total number of steps per day (filled data)",xlab="Number of steps",col="grey")
```


The mean and median total number of steps taken per day.

```{r, echo=TRUE}
new_mean<-mean(totalstepsperday1)
new_median<-median(totalstepsperday1)
```

New mean steps taken per day is **`r new_mean`** (compare to `r mean_steps`)

New median steps taken per day is **`r new_median`** (compare to `r median_steps`)

####Differences in activity patterns between weekdays and weekends

Create a new factor variable in the dataset with two levels – “weekday”
and “weekend”

```{r echo=TRUE}
activity1$day<-as.factor(weekdays(activity1$date,abbreviate=T))
levels(activity1$day)[!(levels(activity1$day)=="Ñá"|levels(activity1$day)=="Íä")] <- "weekday"   
levels(activity1$day)[(levels(activity1$day)=="Ñá"|levels(activity1$day)=="Íä")] <- "weekend"
```

Make a panel plot containing a time series plot of the
5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

```{r}
library(ggplot2)
qplot(interval,steps, data=activity1,stat="summary", fun.y=mean, geom="line")+facet_grid(day~.,scales = "free_y")
```

